\vspace{-3em} \textbf{Occultation mapping of Io's surface in the near-infrared \\ I: Inferring static maps}

### 2021-03-04 paper/paper.tex
Dan's comments

% Title
\title{
    \vspace{-3em}
\textbf{Occultation mapping of Io's surface in the near-infrared \\ I: Inferring static maps} 
}
% Author list
\author{Fran Bartoli\'c}
\email{fb90@st-andrews.ac.uk}
\affil{Centre~for~Exoplanet~Science, SUPA, School~of~Physics~and~Astronomy, University~of~St.~Andrews, St.~Andrews, UK}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY, USA}
\author{Rodrigo Luger}
\author{Daniel Foreman-Mackey}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY, USA}
\author{Robert R. Howell}
\affil{Department~of~Geology~\&~Geophysics, University~of~Wyoming, Laramie, WY, USA}
\author{Julie A. Rathbun}
\affil{Planetary~Science~Institute, Tuscon, AZ, USA}
%

\begin{abstract}
Jupiter's moon Io is the most volcanically active body in the Solar System with hundreds of active volcanoes varying in intensity on different timescales.
Existing photometric observations in the near infrared taken during occultations by Jupiter and other Galilean moons encode a wealth of information about its surface.
These data are ideal for testing methods for inferring surface features of occulted bodies from time series observations.
We built a generative model of Io's surface using the code starry \href{https://rodluger.github.io/starry/}{\color{linkcolor}\faGithub} which enables fast analytic and differentiable computation of occultation light curves in emitted and reflected light.
    Our fully Bayesian model is able to recover known bright features on the surface of Io using only two light curves and without any assumptions on the locations, shapes or the number of features.
    The methods we have developed are also directly applicable to the problem of mapping stars and exoplanets. \href{https://github.com/fbartolic/volcano}{\color{linkcolor}\faGithub}
\end{abstract}

%
\section{Introduction}
The surface of Io is covered with hundreds of volcanoes which appear as bright spots in the near infrared, their intensities varying on timescales ranging from days to decades.
The global heat flow on its surface is about 40 times larger than Earth's \citep{breuer2007,davies2010} and about 50\% of the heat flow emanates from only 1.2\% of Io's surface \citep{veeder2012}.
This intense volcanic activity cannot be explained by radioactive decay, the mechanism which drives Earth's volcanism; instead, the volcanism on Io is driven by tidal interactions with Jupiter and sustained by the Laplace resonance with Europa and Ganymede \citep{peale1979}.
Contrary to Earth, Io lacks plate tectonics.
The heat transport mechanism likely operating on Io is volcanism through the lithosphere (the outermost shell of a terrestrial body), also known as \emph{Heat-pipe volcanism}, a process in which magma is transported to the surface through localised vents \citep{oreilly1981}.
Heat-pipe volcanism might have occurred in the early history of terrestrial planets in the Solar System \citep{moore2013,moore2017}, most notably on early Earth prior to the onset of plate tectonics.

Besides providing a window into early volcanic activity in the Solar System, Io is also in many ways an analogue of a volcanically active exoplanet.
Volcanic exoplanets, sometimes dubbed \emph{super-Ios} or \emph{lava worlds}, have gathered a lot of interest in recent years because both photometric and spectral signatures of volcanic activity on such worlds will likely be detectable in the near future with telescopes such as JWST and LUVOIR \citep{kaltenegger2010,henning2018,oza2019,chao2020}.
Existing exoplanet detections of planets with potential volcanic activity include CoRoT-7b \citep{barnes2010}, the first rocky exoplanet discovered and one which is likely heated by strong tidal forces; 55 Cancri e, whose inferred longitudinal offset in peak surface emission has been attributed to, among other things, lava flows on the surface \citep{demory2016,demory2016d,hammond2017}; several planets in the TRAPPIST-1 system in a Laplace-like resonance with likely volcanic activity \citep{kislyakova2017,dobos2019}, and many others.
Not all such exoplanets will have Io-like volcanism; some may have magma oceans due to their proximity to the star or volcanism caused by nuclear decay, similar to volcanism on present day Earth.

@@

Most studies of occultations of Io fit multiple light curves independently with the goal of inferring one-dimensional longitudinal variations in brightness on its surface (notable exceptions are \cite{spencer1994} and \cite{dekleer2017}), assuming strong priors on the locations of individual volcanoes based on maps from space mission observations.
In this work, we build a fully probabilistic model in order to infer a \emph{two-dimensional map} of Io's volcanic surface from archival IRTF observations of Io in the near infrared.
We assume that the map corresponding to a given set of light curves is static (up to an overall amplitude), that is, that it didn't change substantially between the times of observations.
In Paper II in this series we will relax this assumption by constructing a model which can be used to infer a time-variable map of Io from an any number of light curves observed at arbitrary times.

@@

Our main goal in this paper is to develop a general model for mapping surfaces of occulted bodies with sparse, high contrast features and apply it to observations of Io.
Because we have lots of prior information on what the surface of Io should look like, we can compare features in our inferred maps to known volcanic hotspots and test the sensitivity of the inferred maps to different choices of priors and noise models.
The paper is organized as follows.
In \S\ref{sec:data} we describe the IRTF light curves of occultations by Jupiter which we use to infer maps.
In \S\ref{sec:model} we discuss the generative model for the data and we write down the likelihood function.
In \S\ref{sec:inverse_problem} we focus on the question of how to do Bayesian inference given the forward model specifications defined in \S\ref{sec:model}.
We discuss and test different priors on map features, we quantify the information content of light curves and fit the models using Hamiltonian Monte Carlo (HMC). 
We then test the model on realistic simulated data.
In \S\ref{sec:results} we show the results on real IRTF data for two pairs of light curves, one pair observed in 1998 and another in 2017.
We show the inferred maps and the parameters quantifying the location and intensity of the hot spots, we discuss the sensitivity of the results to different choices of priors and plot the inferred hotspots on an optical map of Io constructed from Galieleo observations.
Finally, in \S\ref{sec:discussion} we summarize the main results of the paper and discuss potential applications of our model to observations of volcanic exoplanets.

@@

All of the figures in this paper were auto-generated
using the Azure Pipelines continuous integration (CI) service.
Icons next to each of the figures \codeicon \,
link to the exact script used to generate them to ensure the reproducibility
of our results.

@@

\begin{figure}[h!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/irtf_max_flux.pdf}
    \oscaption{irtf_dataset}{%
        Maximum flux for each IRTF light curve of Io observed during an occultation of Io by Jupiter.
       \label{fig:irtf_max_flux}
    }
    \end{centering}
\end{figure}

The complete dataset consists of 112 sets of observations of Io in the near infrared taken from 1996 until 2018 using different instruments at NASA's IRTF observatory, the NSFCam  1-5 $\mu m$ camera \citep{shure1994}, the SpeX 0.8-5.5 $\mu m$ spectrographer/imager \citep{rayner2003} and the iShell 1.1-5.3 $\mu m$  spectrographer/imager \citep{johnrayner2016}.
%(TODO: fix citation).
Although the observations span decades, the timespan between observations varies and there were very few observations taken between 2008 and 2016.
All of the observations were taken while Io was \emph{in eclipse}, meaning that it was in Jupiter's shadow so the source of observed emission from the surface is thermal radiation.
On the other hand, for observations of Io \emph{in sunlight} the observed flux is a sum of thermal (volcanic) emission and reflected sunlight which depends on the albedo map in the near infrared.
In figure~\ref{fig:irtf_max_flux} we plot the maximum flux for each occultation light curve in units of $\mathrm{GW}/\mu \mathrm{m}/\mathrm{sr}$ as a function of time.
This value is approximately equal to the total flux emitted from the Jupiter-facing side of Io.
The major variation in the baseline brightness is driven by Loki.
Most hotspots on Io can be divided into those who are persistently active for a year or longer at moderate intensity, and those which have bursts of activity and reach very high intensities for short periods of time \citep{dekleer2016}.

@@

\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=1.\linewidth]{figures/pixels_to_spherical_harmonics.pdf}
    \oscaption{pixels_to_harmonics}{%
        Illustration of the transformation from a pixel map (left) defined on a fixed grid on the sphere into the spherical harmonic basis at $l=20$ (right) via a linear operator $\mathbf{P}^\dagger$.
        The intensity of each pixel was drawn independently from an exponential prior.
        The histograms below each of the maps show the distributions of intensities on the map. 
        Since the pixel map is higher resolution than the spherical harmonic map at $l=20$, the         spherical harmonic map appears smoother and the intensity distribution is more similar to a skewed Gaussian rather than an exponential distribution.
        Despite this fact we find that fitting for maps in the pixel basis and transforming them to spherical harmonics at each MCMC step in order to compute the light curve analytically with \textsf{starry} is superior to fitting for spherical harmonics because the pixel grid provides a strong constraint on the spherical harmonic map structure.
  \label{fig:pixels_to_harmonics}}
    \end{centering}
\end{figure}

@@

To model the data covariance $\boldsymbol{\Sigma}$ we use use a Gaussian Process and we compute the likelihood using the fast Celerite method \citep{foreman-mackey2017} as implemented in the \textsf{celerite2} package \citep{foreman-mackey2017a,foreman-mackey2018}.
We use the simple (approximate) Mat\'ern 3/2 kernel function which is parametrized by two values, a standard deviation parameter $\sigma_\mathrm{GP}$ and a characteristic timescale parameter $\rho_\mathrm{GP}$.
It is given by 
\begin{equation}
    k(\tau;\sigma_\mathrm{GP},
    rho_\mathrm{GP})=\sigma_\mathrm{GP}^{2}\left[(1+1 / \epsilon) e^{-(1-\epsilon) \sqrt{3} \tau / \rho_\mathrm{GP}}(1-1 / \epsilon) e^{-(1+\epsilon) \sqrt{3} \tau / \rho_\mathrm{GP}}\right]
    \quad,
\end{equation}
where $\tau=|t_n-t_m|$ and $\epsilon$ controls the quality of the approximation.
A single element of the data covariance matrix $\boldsymbol{\Sigma}$ is then 
\begin{equation}
    \boldsymbol{\Sigma}_{nm}=\sigma_n^2\delta_{nm} + k(\tau;\sigma_\mathrm{GP},\rho_\mathrm{GP})
    \quad,
    \label{eq:data_covariance_element}
\end{equation}
where $\sigma_n$ is the errorbar for the n-th data point.
Since we fit multiple independent light curves the total log likelihood is the sum of individual likelihoods defined in Equation~(\ref{eq:likelihood}).

@@

The total log likelihood is the sum of the log likelihoods for individual light curves (Equation~(\ref{eq:likelihood})). 
To compute it we also need to specify the data covariance matrix defined in Equation~(\ref{eq:data_covariance_element}).
Each element of the covariance matrix consists of a white noise component (errobars) and a correlated noise component (Gaussian Process).
Since the IRTF light curves are not provided with estimated errobars, we choose to fit for all errobars simultaneously using a hierarchical approach in which we assign a global scale $l$ for all errorbars in a given light curve and we draw each individual errobar from a Half Normal distribution with standard deviation equal to that scale.
That is, we have 
\begin{align}
    \sigma_{i,I}&\sim \mathcal{N}^+(0, l_E),\quad i=1,\dots,N_I\\
    \sigma_{i,E}&\sim \mathcal{N}^+(0,l_I),\quad i=1,\dots, N_E
\end{align}
where $\sigma_{i,I}$ are the errobars for the ingress light curve and $\sigma_{i,E}$ are the errobars for the egress light curve.
The symbol $\mathcal{N}^+$ denotes a truncated Gaussian distribution restricted to positive values.
Treating errobars as free parameters means that we are introducing hundreds of more parameters (as many as there are data points) which is not an issue as long as these parameters are constrained by the data and our model is regularized.
The noise model as defined above is extremely flexible, it can account for a given feature in the light curve by inflating individual errobars (white noise) or by varying the characteristic timescale $\rho_\mathrm{GP}$ in the Matern 3/2 kernel.
All model parameters and their associated priors are listed in Table~\ref{tab:priors}.

@@

To see how the above results change if we don't include a Gaussian Process in the noise model we fit the same pair of occultations from 1998 using the same priors, the results are shown in Figure~\ref{fig:irtf_1998_no_GP}.
The main difference in the inferred map compared to Figure~\ref{fig:irtf_1998} is that there are two extra spots visible in the map.
The two spots in the eastern hemisphere are the result of the model struggling to explain the main feature in the light curve at around 3.2 minutes in ingress light curve and 3.8 minutes in the egress light curve, which is due to  Loki coming out and into view during the occultation.
The model inflates the errobars around those times because it unable to make the spot small enough.
This results in the single hot spot at the location of Loki shown in Figure~\ref{fig:irtf_1998} morphing into two spots.
While this feature is almost certainly spurious, the other spot in the northwestern hemisphere appears more likely to be real.
Looking at the miniature maps in Figure~\ref{fig:irtf_1998_no_GP}, this spot is in view only at egress and it corresponds to a small step in the light curve starting at 0.9 minutes.
The estimated location of this spot is $20.449_{-0.066}^{+1.253}\quad^\circ$ N latitude and $53.107_{-0.115}^{+0.296}\quad^\circ$ W longitude.
This location corresponds to the mountain Mongibello Mons but it does not correspond to a well known hotspot.
Overall, we conclude that including the Gaussian Process prevents the physical model from overfitting the data although it might occasionally pick up a feature which is due to a faint real hot spot.
With a particular scientific goal in mind, it is straightforward to experiment with different noise models to test how the properties of any given spot change with different assumptions.

@@

\section{Discussion}
\label{sec:discussion}
\subsection{Occultation mapping of Io}
We have presented a novel method for mapping the volcanic emission on Io from occultation light curves.
The method relies on the \textsf{starry} algorithm which enables fast analytic computation of occultation light curves by expanding the surface emission map in spherical harmonics.
Our method is substantially different from past work because we do not assume where the volcanic features are located on the surface of Io, how many there are or what they look like.
Instead, we place weaker assumptions on the global structure of the map by requiring that the inferred map has positive intensity everywhere and most importantly, that it is \emph{sparse}.
\cite{aizawa2020} also find the sparsity assumption useful in the context of exoplanet mapping.
Besides the model for the surface map, our method also incorporates a sophisticated noise model which includes a Gaussian Process and a hierarchical model for the (unknown) errorbars.
As a result of the sparsity and positivity constraints and the flexibility of the noise model,
our model is parsimonious: it places features on the map only if the data provides strong evidence for the existence of those features.
This property is not always desirable but it works very well for Io, whose surface is covered with small highly localized and bright volcanic hotspots.
Our model is superior to parametric methods which assume some fixed number of spots on the surface of Io and parametrize the spot properties because we do not need to make such strong assumptions about the surface and the computational cost of the parametric models is comparable to our approach.

@@

The main limitation of our model besides the fact that we cannot account for time-dependent maps is the limited resolution of the spherical harmonic maps for which \textsf{starry} can compute occultation light curves.
Although we can still constrain the peak intensity of hotspots (or some other measure of the center of emission) with much higher precision, we lose the ability to resolve two spots close to each other and the ability to constrain the actual size of the spots. 
It is quite possible that there is a way around this problem if we compute the various integrals in \textsf{starry} numerically using high precision arithmetic.
Since all these operations are contained in the design matrix $\mathbf{A}$ (assuming that the ephemeris is fixed) we would only need to do the expensive computation once for each occultation.
To be able to constrain the size of the Loki hotspot for example, we would need to fit maps on the order of $l\approx 50$.
However, the kind of detailed mapping of Loki's magma lake with a precision of a few km done by \cite{dekleer2017} will not be possible with a spherical harmonic model because it would require maps of extremely large degrees ($l\gtrapprox 700$).

@@

Besides being useful for modeling secondary eclipses of volcanic Super-Earths, our model is also likely to useful for modeling features on gaseous exoplanets with sparse surface features.
These will be much easier to observe with JWST  than Super-Earths because of their larger size and higher surface temperatures.
Even if the sparsity assumption is not valid for these types of planets the hybrid pixel/spherical harmonic model can be easily be used with a different set of priors more appropriate for observations.
For example, a sensible prior for the surface of a fast rotating gaseous exoplanet might be requiring some degree of azimuthal symmetry.

@@

When using the Regularized Horseshoe prior in a small data regime it is often necessary to use the non-centered parametrization to avoid funnels in the posterior which are often present in hierarchical models\footnote{\url{https://mc-stan.org/docs/2_26/stan-users-guide/reparameterization-section.html}}.
The purpose of this reparametrization is to reduce the dependence between the hyperparameters in the posterior.
To implement the non-centered parametrization we replace priors in Equation~\ref{eq:reg_horseshoe} with zero mean and unit variance priors and rescale them with deterministic transforms as follows
\begin{equation}
\begin{aligned}
    &\overline{\tau} \sim \mathrm{Half}-\mathcal{C}\left(0, 1\right)\\
    &\overline{c^{2}}  \sim \mathrm{Inv}-\mathcal{G}\left(\frac{\nu}{2}, 1\right) \\
    &\overline{\lambda}_{i}  \sim \mathrm{Half}-\mathcal{C}(0,1) \\
    &\overline{p}_{i}  \sim \mathcal{N}^+\left(0, 1\right)\\
    &\tau=\tau_0\overline{\tau}\\
    &c^2=\frac{\nu}{2}s^2\,\overline{c^2}\\
    &\lambda_{i} =\frac{c \overline{\lambda}_{i}}{\sqrt{c^{2}+\tau^{2} \overline{\lambda}_{i}^{2}}} \\
    &p_i = \tau\lambda_i\,\overline{p}_i
\end{aligned}
    \label{eq:reg_horseshoe_noncentered}
\end{equation}
We find that without using the non-centered parametrization the sampling is problematic and there are lots of divergences in the gradients of the parameters; with the non-centered parametrization there are no problems with sampling.

>>>>

% Title
\title{
    \vspace{-3em}
\textbf{Occultation mapping of Io's surface in the near-infrared \\ I: Inferring static maps} 
}
% Author list
\author[0000-0001-8630-9794]{Fran Bartoli\'c}
\email{fb90@st-andrews.ac.uk}
\affil{Centre~for~Exoplanet~Science, SUPA, School~of~Physics~and~Astronomy, University~of~St.~Andrews, St.~Andrews, UK}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY, USA}
\author[0000-0002-0296-3826]{Rodrigo Luger}
\author[0000-0002-9328-5652]{Daniel Foreman-Mackey}
\affil{Center~for~Computational~Astrophysics, Flatiron~Institute, New~York, NY, USA}
\author[0000-0003-4859-2060]{Robert R. Howell}
\affil{Department~of~Geology~\&~Geophysics, University~of~Wyoming, Laramie, WY, USA}
\author[0000-0001-7619-652X]{Julie A. Rathbun}
\affil{Planetary~Science~Institute, Tuscon, AZ, USA}
%

\begin{abstract}
Jupiter's moon Io is the most volcanically active body in the Solar System with hundreds of active volcanoes varying in intensity on different timescales.
Existing photometric observations in the near infrared taken during occultations by Jupiter and other Galilean moons encode a wealth of information about its surface.
These data are ideal for testing methods for inferring surface features of occulted bodies from time series observations.
We built a generative model of Io's surface using the code starry \href{https://rodluger.github.io/starry/}{\color{linkcolor}\faGithub} which enables fast analytic and differentiable computation of occultation light curves in emitted and reflected light.
    Our probabilistic Bayesian model is able to recover known bright features on the surface of Io using only two light curves and without any assumptions on the locations, shapes or the number of features.
    The methods we have developed are also directly applicable to the problem of mapping stars and exoplanets. \href{https://github.com/fbartolic/volcano}{\color{linkcolor}\faGithub}
\end{abstract}

%
\section{Introduction}
The surface of Io is covered with hundreds of volcanoes which appear as bright spots in the near infrared, their intensities varying on timescales ranging from days to decades.
The global heat flow on its surface is about 40 times larger than Earth's \citep{breuer2007,davies2010} and about 50\% of the heat flow emanates from only 1.2\% of Io's surface \citep{veeder2012}.
This intense volcanic activity cannot be explained by radioactive decay, the mechanism which drives Earth's volcanism; instead, the volcanism on Io is driven by tidal interactions with Jupiter and sustained by the Laplace resonance with Europa and Ganymede \citep{peale1979}.
Contrary to Earth, Io lacks plate tectonics.
The heat transport mechanism likely operating on Io is volcanism through the lithosphere (the outermost shell of a terrestrial body), also known as \emph{heat-pipe volcanism}, a process in which magma is transported to the surface through localised vents \citep{oreilly1981}.
Heat-pipe volcanism might have occurred in the early history of terrestrial planets in the Solar System \citep{moore2013,moore2017}, most notably on early Earth prior to the onset of plate tectonics.

Besides providing a window into early volcanic activity in the Solar System, Io is also in many ways an analogue of a volcanically active exoplanet.
Volcanic exoplanets, sometimes dubbed \emph{super-Ios} or \emph{lava worlds}, have gathered a lot of interest in recent years because both photometric and spectral signatures of volcanic activity on such worlds will likely be detectable in the near future with telescopes such as JWST and LUVOIR \citep{kaltenegger2010,henning2018,oza2019,chao2020}.
Existing exoplanet detections of planets with potential volcanic activity include CoRoT-7b \citep{barnes2010}, the first rocky exoplanet discovered and one which is likely heated by strong tidal forces; 55 Cancri e, whose inferred longitudinal offset in peak surface emission has been attributed to, among other things, lava flows on the surface \citep{demory2016,demory2016d,hammond2017a}; several planets in the TRAPPIST-1 system in a Laplace-like resonance with likely volcanic activity \citep{kislyakova2017,dobos2019}, and many others.
Not all such exoplanets will have Io-like volcanism; some may have magma oceans due to their proximity to the star or volcanism caused by nuclear decay, similar to volcanism on present day Earth.

@@

Most studies of (unresolved) occultations of Io fit multiple light curves independently with the goal of inferring one-dimensional longitudinal variations in brightness on its surface (notable exceptions are \cite{spencer1994} and \cite{dekleer2017}), assuming relatively strong priors on the locations of individual volcanoes.
In this work, we build a fully probabilistic model in order to infer a \emph{two-dimensional map} of Io's volcanic surface from archival IRTF observations of Io in the near infrared.
We assume that the map corresponding to a given set of light curves is static (up to an overall amplitude), that is, that it didn't change substantially between the times of observations.
In Paper II in this series we will relax this assumption by constructing a model which can be used to infer a time-variable map of Io from an any number of light curves observed at arbitrary times.

@@

Our main goal in this paper is to develop a general model for mapping surfaces of occulted bodies with sparse, high contrast features and apply it to observations of Io.
Because of high resolution resolved imaging of Io, we can compare features in our inferred maps to known volcanic hotspots and test the sensitivity of the inferred maps to different choices of priors and noise models.
The paper is organized as follows.
In \S\ref{sec:data} we describe the IRTF light curves of occultations by Jupiter which we use to infer maps.
In \S\ref{sec:model} we discuss the generative model for the data and we write down the likelihood function.
In \S\ref{sec:inverse_problem} we focus on the question of how to do Bayesian inference given the forward model specifications defined in \S\ref{sec:model}.
We discuss and test different priors on map features, we quantify the information content of light curves and fit the models using Hamiltonian Monte Carlo (HMC). 
We then test the model on realistic simulated data.
In \S\ref{sec:results} we show the results on real IRTF data for two pairs of light curves, one pair observed in 1998 and another in 2017.
We show the inferred maps and the parameters quantifying the location and intensity of the hot spots, we discuss the sensitivity of the results to different choices of priors and plot the inferred hotspots on an optical map of Io constructed from Galieleo observations.
Finally, in \S\ref{sec:discussion} we summarize the main results of the paper and discuss potential applications of our model to observations of volcanic exoplanets.

@@

The code associated with this paper is publicly hosted in a GitHub repository\footnote{\url{https://github.com/fbartolic/volcano}}. 
All of the figures in this paper were auto-generated
using the Azure Pipelines continuous integration (CI) service.
Icons next to each of the figures \codeicon \,
link to the exact script used to generate them to ensure the reproducibility
of our results.

@@

\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=\linewidth]{figures/irtf_max_flux.pdf}
    \oscaption{irtf_dataset}{%
        Maximum flux for each IRTF light curve of Io observed during an occultation of Io by Jupiter.
       \label{fig:irtf_max_flux}
    }
    \end{centering}
\end{figure}

The complete dataset consists of 112 sets of observations of Io in the near infrared taken from 1996 until 2018 using different instruments at NASA's IRTF observatory, the NSFCam  1-5 $\mu m$ camera \citep{shure1994}, the SpeX 0.8-5.5 $\mu m$ spectrographer/imager \citep{rayner2003} and the iShell 1.1-5.3 $\mu m$  spectrographer/imager \citep{rayner2016}.
Although the observations span decades, the timespan between observations varies and there were very few observations taken between 2008 and 2016.
All of the observations were taken while Io was \emph{in eclipse}, meaning that it was in Jupiter's shadow so the source of observed emission from the surface is thermal radiation.
On the other hand, for observations of Io \emph{in sunlight} the observed flux is a sum of thermal (volcanic) emission and reflected sunlight which depends on the albedo map in the near infrared.
In figure~\ref{fig:irtf_max_flux} we plot the maximum flux for each occultation light curve in units of $\mathrm{GW}/\mu \mathrm{m}/\mathrm{sr}$ as a function of time.
This value is approximately equal to the total flux emitted from the Jupiter-facing side of Io.
The major variation in the baseline brightness is driven by Loki.
Most hotspots on Io can be divided into those who are persistently active for a year or longer at moderate intensity, and those which have bursts of activity and reach very high intensities for short periods of time \citep{dekleer2016}.

@@

\begin{figure}[t!]
    \begin{centering}
    \includegraphics[width=1.\linewidth]{figures/pixels_to_spherical_harmonics.pdf}
    \oscaption{pixels_to_harmonics}{%
        Illustration of the transformation from a pixel map (left) defined on a fixed grid on the sphere into the spherical harmonic basis at $l=20$ (right) via a linear operator $\mathbf{P}^\dagger$.
        The intensity of each pixel was drawn independently from an exponential prior.
        The histograms below each of the maps show the distributions of intensities on the map. 
        Since the pixel map is higher resolution than the spherical harmonic map at $l=20$, the         spherical harmonic map appears smoother and the intensity distribution is more similar to a skewed Gaussian rather than an exponential distribution.
        Despite this fact, we find that fitting for maps in the pixel basis and transforming them to spherical harmonics at each MCMC step in order to compute the light curve analytically with \textsf{starry} is better than fitting for spherical harmonics because the pixel grid provides a strong constraint on the spherical harmonic map structure.
  \label{fig:pixels_to_harmonics}}
    \end{centering}
\end{figure}

@@

To model the data covariance $\boldsymbol{\Sigma}$ we use use a Gaussian Process and we compute the likelihood using the fast Celerite method \citep{foreman-mackey2017} as implemented in the \textsf{celerite2} package \citep{foreman-mackey2017a,foreman-mackey2018}.
We use the simple (approximate) Mat\'ern 3/2 kernel function which is parametrized by two values, a standard deviation parameter $\sigma_\mathrm{GP}$ and a characteristic timescale parameter $\rho_\mathrm{GP}$.
The Mat\'ern 3/2 kernel is defined by
\begin{equation}
    k(\tau;\sigma_\mathrm{GP},
    \rho_\mathrm{GP})=\
    \sigma_\mathrm{GP}^{2}\left(1+\frac{\sqrt{3} \tau}{\rho_\mathrm{GP}}\right) \exp \left(-\frac{\sqrt{3} \tau}{\rho_\mathrm{GP}}\right)
    \quad,
\end{equation}
where $\tau=|t_n-t_m|$ and $\epsilon$ controls the quality of the approximation.
A single element of the data covariance matrix $\boldsymbol{\Sigma}$ is then 
\begin{equation}
    \boldsymbol{\Sigma}_{nm}=\sigma_n^2\delta_{nm} + k(\tau;\sigma_\mathrm{GP},\rho_\mathrm{GP})
    \quad,
    \label{eq:data_covariance_element}
\end{equation}
where $\sigma_n$ is the errorbar for the n-th data point.
Since we fit multiple independent light curves the total log likelihood is the sum of individual likelihoods defined in Equation~(\ref{eq:likelihood}).

@@

The total log likelihood is the sum of the log likelihoods for individual light curves (Equation~(\ref{eq:likelihood})). 
To compute it we also need to specify the data covariance matrix defined in Equation~(\ref{eq:data_covariance_element}).
Each element of the covariance matrix consists of a white noise component (errobars) and a correlated noise component (Gaussian Process).
Since the IRTF light curves are not provided with estimated errobars, we choose to fit for all errobars simultaneously using a hierarchical approach in which we assign a global scale $l$ for all errorbars in a given light curve and we draw each individual errobar from a Half Normal distribution with standard deviation equal to that scale.
That is, we have 
\begin{align}
    \sigma_{i,I}&\sim \mathcal{N}^+(0, l_E),\quad i=1,\dots,N_I\\
    \sigma_{i,E}&\sim \mathcal{N}^+(0,l_I),\quad i=1,\dots, N_E
\end{align}
where $\sigma_{i,I}$ are the errobars for the ingress light curve and $\sigma_{i,E}$ are the errobars for the egress light curve.
The symbol $\mathcal{N}^+$ denotes a truncated Gaussian distribution restricted to positive values.
Treating errobars as free parameters means that we are introducing hundreds of more parameters (as many as there are data points) which is not an issue as long as these parameters are constrained by the data and our model is regularized.
The noise model as defined above is extremely flexible, it can account for a given feature in the light curve by inflating individual errobars (white noise) or by varying the characteristic timescale $\rho_\mathrm{GP}$ in the Mat\'ern 3/2 kernel.
All model parameters and their associated priors are listed in Table~\ref{tab:priors}.

@@

To see how the above results change if we don't include a Gaussian Process in the noise model we fit the same pair of occultations from 1998 using the same priors, the results are shown in Figure~\ref{fig:irtf_1998_no_GP}.
The main difference in the inferred map compared to Figure~\ref{fig:irtf_1998} is that there are two extra spots visible in the map.
The two spots in the eastern hemisphere are the result of the model struggling to explain the main feature in the light curve at around 3.2 minutes in ingress light curve and 3.8 minutes in the egress light curve, which is due to  Loki coming out and into view during the occultation.
The model inflates the errobars around those times because it unable to make the spot small enough.
This results in the single hot spot at the location of Loki shown in Figure~\ref{fig:irtf_1998} morphing into two spots.
While this feature is almost certainly spurious, the other spot in the northwestern hemisphere appears more likely to be real.
Looking at the miniature maps in Figure~\ref{fig:irtf_1998_no_GP}, this spot is in view only at egress and it corresponds to a small step in the light curve starting at 0.9 minutes.
The estimated location of this spot is approximatel $\sim 20\quad^\circ$ N latitude and $\sim 69\quad^\circ$ W longitude.
This location corresponds to southern end of the mountain Mongibello Mons but there are no known persistent hotspots at that location so we cannot say if the hotspots is real or not without independently detecting it in other light curves.
Overall, we conclude that including the Gaussian Process prevents the physical model from overfitting the data although it might occasionally pick up a feature which is due to a faint real hot spot.
With a particular scientific goal in mind, it is straightforward to experiment with different noise models to test how the properties of any given spot change with different assumptions.

@@

\section{Discussion}
\label{sec:discussion}
\subsection{Occultation mapping of Io}
We have presented a novel method for mapping the volcanic emission on Io from occultation light curves.
The method relies on the \textsf{starry} algorithm which enables fast analytic computation of occultation light curves by expanding the surface emission map in spherical harmonics.
Our method is substantially different from past work because we do not assume where the volcanic features are located on the surface of Io, how many there are or what they look like.
Instead, we place weaker assumptions on the global structure of the map by requiring that the inferred map has positive intensity everywhere and most importantly, that it is \emph{sparse}.
\cite{aizawa2020} also find the sparsity assumption useful in the context of exoplanet mapping.
Besides the model for the surface map, our method also incorporates a sophisticated noise model which includes a Gaussian Process and a hierarchical model for the (unknown) errorbars.
As a result of the sparsity and positivity constraints and the flexibility of the noise model,
our model is parsimonious: it places features on the map only if the data provides strong evidence for the existence of those features.
This property is not always desirable but it works very well for Io, whose surface is covered with small highly localized and bright volcanic hotspots.
Our model is better than parametric methods which assume some fixed number of spots on the surface of Io and parametrize the spot properties because we do not need to make such strong assumptions about the surface and the computational cost of the parametric models is comparable to our approach.

@@

The main limitation of our model besides the fact that we do not yet account for time-dependent maps is the limited resolution of the spherical harmonic maps for which \textsf{starry} can compute occultation light curves.
Although we can still constrain the peak intensity of hotspots (or some other measure of the center of emission) with much higher precision, we lose the ability to resolve two spots close to each other and the ability to constrain the actual size of the spots. 
It is quite possible that there is a way around this problem if we compute the various integrals in \textsf{starry} numerically using high precision arithmetic.
Since all these operations are contained in the design matrix $\mathbf{A}$ (assuming that the ephemeris is fixed) we would only need to do the expensive computation once for each occultation.
To be able to constrain the size of the Loki hotspot for example, we would need to fit maps on the order of $l\approx 50$.
However, the kind of detailed mapping of Loki's magma lake with a precision of a few km done by \cite{dekleer2017} will not be possible with a spherical harmonic model because it would require maps of extremely large degrees ($l\gtrapprox 700$).

@@

Besides being useful for modeling secondary eclipses of volcanic Super-Earths, our model is also likely to be useful for modeling features on gaseous exoplanets with sparse surface features.
These will be much easier to observe with JWST  than Super-Earths because of their larger size and higher surface temperatures.
Even if the sparsity assumption is not valid for these types of planets the hybrid pixel/spherical harmonic model can be easily be used with a different set of priors more appropriate for observations.
For example, a sensible prior for the surface of a fast rotating gaseous exoplanet might be requiring some degree of azimuthal symmetry.

@@

When using the Regularized Horseshoe prior in a small data regime it is often necessary to use the non-centered parametrization to avoid funnels in the posterior which are often present in hierarchical models\footnote{\url{https://mc-stan.org/docs/2_26/stan-users-guide/reparameterization-section.html}}.
The purpose of this reparametrization is to reduce the dependence between the hyperparameters in the posterior.
To implement the non-centered parametrization we replace priors in Equation~\ref{eq:reg_horseshoe} with zero mean and unit variance priors and rescale them with deterministic transforms as follows
\begin{equation}
\begin{aligned}
    &\overline{\tau} \sim \mathrm{Half}-\mathcal{C}\left(0, 1\right)\\
    &\overline{c^{2}}  \sim \mathrm{Inv}-\mathcal{G}\left(\frac{\nu}{2}, 1\right) \\
    &\overline{\lambda}_{i}  \sim \mathrm{Half}-\mathcal{C}(0,1) \\
    &\overline{p}_{i}  \sim \mathcal{N}^+\left(0, 1\right)\\
    &\tau=\tau_0\overline{\tau}\\
    &c^2=\frac{\nu}{2}s^2\,\overline{c^2}\\
    &\lambda_{i} =\frac{c \overline{\lambda}_{i}}{\sqrt{c^{2}+\tau^{2} \overline{\lambda}_{i}^{2}}} \\
    &p_i = \tau\lambda_i\,\overline{p}_i
\end{aligned}
    \label{eq:reg_horseshoe_noncentered}
\end{equation}
We find that without using the non-centered parametrization the sampling is problematic and there are many divergences in the gradients of the parameters; with the non-centered parametrization there are no problems with sampling.
